---
title: "Reproducible research course project 1"
output: html_document
---

```{r download file, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Ex <- read.csv("activity.csv")
names(Ex)
str(Ex)
Ex$date <- as.Date(Ex$date)
```

```{r Histogram of total daily steps, echo = TRUE}
TotalDailySteps<-with(Ex, tapply(steps, date, sum, na.rm=TRUE))
hist(TotalDailySteps, col = "maroon")
```
The next step is to report the mean and median of total number of steps taken each day. This is a simple and straight forward task
```{r mean and media, echo=TRUE}
meanStep<-round(mean(TotalDailySteps), 2)
medianStep<-round(median(TotalDailySteps), 2)
```
 so the meanstep is 9354.23 and medianStep is 10395
 
#What is the average daily activity pattern?
For this part of analysis, we should calculate average steps taken on each 15 minutes interval across all study period. Hence, I calculated it into a new variable.
 
``` {r interval seps, echo=TRUE}
intervalSteps<-with(Ex, tapply(steps, interval, mean, na.rm=TRUE)) 
```
Now, we can plot the mean step value for each interval. Again, 'base plotting system' is more the sufficient for this step.

```{r baseplot, echo=TRUE}
plot(intervalSteps,axes = F, type="l", col="maroon", xlab="Time", ylab="Average Number of Steps", main="Average Daily Activity Pattern")
axis(1,at=c(0, 36, 72, 108, 144, 180, 216, 252, 288), label = c("0:00", "3:00","6:00", "9:00", "12:00","15:00","18:00","21:00","24:00"))
axis(2)
```
We also have been asked about the interval with maximum average steps taken within. We can calculate it as follow:

```{r interval steps, echo=TRUE}
intervalSteps[which.max(intervalSteps)]
```


#Imputing missing values
First, I want to report the number of missing values.

``` {r missing values, echo=TRUE}
NAcount<-sum(is.na(Ex$steps))
NAcount
```
In this data frame, there are 2304 rows with missing value for 'Steps' variable.

#For imputing missing values, I use the very nice "MICE" package.

```{r load MICE package, echo=TRUE}
library(mice)
imputedValues<-mice(Ex)
```
Now we have imputed values and can reconstruct the new database:

```{r, echo=TRUE}
imputedEx<-complete(imputedValues)
```
Preprocessing of new dataframe:

``` {r, echo=TRUE}
imputedEx$date<-as.Date(imputedEx$date)
```
Calculating total daily steps:

```{r, echo=TRUE}
TotalDailySteps_Ex<-with(imputedEx, tapply(steps, date, sum, na.rm=TRUE))
```
Making the histograms:
```{r, echo=TRUE}
hist(TotalDailySteps_Ex, col = "pink")
```

Calculating the central values:
```{r, echo=TRUE}
meanStep_Ex<-round(mean(TotalDailySteps_Ex), 2)
medianStep_Ex<-round(median(TotalDailySteps_Ex), 2)
```
So, after imputation of the missing values, the mean total daily steps is 1.12552610^{4} and the median is 1.127910^{4}.

Calculating the resulted change in central values:
```{r, echo=TRUE}
meanDiff<- meanStep - meanStep_Ex
medianDiff<- medianStep - medianStep_Ex
```
After imputation, change in mean value is -1901.03 and change in median value is -884


##Are there differences in activity patterns between weekdays and weekends?

In the last step, we want to compare the difference between activity on weekdays and weekends.
```{r, echo=TRUE}
imputedEx$weekdays<-weekdays(imputedEx$date)
imputedEx$dayType<-ifelse(imputedEx$weekdays%in%c("Saturday", "Sunday"), "Weekend", "Weekday")
```

Now, we can calculate average steps in each interval based on type of weekday:
```{r, echo=TRUE}
intervalDaySteps<-aggregate(steps~interval+dayType, data = imputedEx, mean)
```

For plotting data in two panels, I use the ggplot2 system:
```{r, echo=TRUE}
library(ggplot2)
g1<-ggplot(intervalDaySteps, aes(interval, steps))
g1 + geom_line() +
        facet_grid(dayType ~ .) +
        xlab("5-minute interval") + 
        ylab("Number of steps")
```

The end!